``Interestingness'' in Mathematical Theory Exploration

### 2019-10-18 PhDThesis/litreview_testing.tex
Wording

Some prominent tools for mathematical theory exploration, like \quickspec{},
were originally developed in the context of \emph{property checking} (although
for \emph{discovering} software properties, rather than checking them), and only
later applied more directly to fields like formal mathematics (e.g. via the
\hipster{} tool for Isabelle/HOL). In particular, the logical domain used by
\quickspec{} (inductive logic with higher-order functions) comes directly from
its application to Haskell; and its search algorithm makes heavy use of the
\quickcheck{} property checker. The boundary between these two fields is hence
rather fuzzy, and cross-pollination of techniques is a compelling target for
research.

Property checking attempts to establish the truth or falsity of a statement (a
``property''), usually involving components of a software system and, crucially,
may contain free variables. \emph{Unit testing} is a common method of software
quality assurance (outside of specialised domains like those which are
life-critical), which can be characterised as proving \emph{existential}
properties: each test includes concrete values for any variables, and checking
is simply a matter of evaluating the code to get a boolean value (pass or fail).
The phrase ``property checking'' is typically reserved for \emph{universal}
properties: those which should hold \emph{for all} values of the variables.
This is a popular method of quality assurance for Haskell software, giving rise
to a number of packages which approach this challenge in different ways.

>>>>

Some prominent tools for mathematical theory exploration, like \quickspec{},
were originally developed in the context of \emph{property checking} (although
for \emph{discovering} software properties, rather than checking them), and only
later applied more directly to fields like formal mathematics (e.g. via the
\hipster{} tool for Isabelle/HOL). In particular, the logical domain used by
\quickspec{} (inductive logic with higher-order functions) comes directly from
its application to Haskell; its search algorithm also makes heavy use of the
\quickcheck{} property checker. The boundary between these two fields is hence
rather fuzzy, and cross-pollination of techniques is a compelling target for
research.

Property checking attempts to establish the truth or falsity of a statement (a
``property''), usually involving components of a software system and, crucially,
which may contain free variables. \emph{Unit testing} is a common method of
software quality assurance (outside of specialised domains like those which are
life-critical), which can be characterised as proving \emph{existential}
properties: each test includes concrete values for any variables, and checking
is simply a matter of evaluating the code to get a boolean value (pass or fail).
The phrase ``property checking'' is typically reserved for \emph{universal}
properties: those which should hold \emph{for all} values of the variables.
This is a popular method of quality assurance for Haskell software, giving rise
to a number of packages which approach this challenge in different ways.
